The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-18 paper.tex
Imported section on novelty effect for previous HRI paper

TDB

>>>>

To illustrate how the novelty of a robot can evoke anthropomorphic effects, we
present an example from our long-term ethnographic study with Roomba vacuum
cleaning robots in households. The observed scenario describes how an elderly
woman (71 years, living on her own with a little dog in a small row house) was
setting up and using the robot for the first time for about 10 minutes.

The woman (E.) is reading aloud parts of the "quick start manual": \emph{"When
Roomba senses dirt, the blue dirt detect light is lit and Roomba cleans more
intensely in that area."} She laughs and turns to the robot, talking directly to
it \emph{"Today you can light up darling!"}. E. carries Roomba to the entrance
door and puts it on the floor [floor type: tiles, partially covered with little
Persian carpets], she explains \emph{"It has to start from here, this is where
you enter the house!"}. After pressing the CLEAN button, the robot quickly gets
stuck with the carpet in the entrance hall, consequently it reports error 1 [
two-tone "Uh-Oh" sound followed by a narrated voice that starts off with "Error
1"]. E. is very surprised \emph{"What, it is even talking to me?"}, lifts the
robot off and puts it back for cleaning. At several occasions, while watching
how the robot moves around and cleans, E. makes compliments to the robot:
\emph{"Nice!"} or \emph{"Good job!"} or \emph{"Bravo!"}. Once, when Roomba turns
right in front of some crumbs on the kitchen floor and moves away without
cleaning them up, E. says: \emph{"There you forgot something, what is with over
there? Look, there you still have something to clean up!"}, and she points at
where she wants Roomba to go. Further, she uses directing gestures and tells the
robot: \emph{"No, go this way! Come over here!"}. The woman keeps observing the
robot and wonders about its "intentions", asking it directly: \emph{"Why do you
leave the kitchen? There is still something there on the fringe!"}. E. turns to
the experimenter, and tells her: \emph{"She [the robot] loves the hall!"}.

Besides this behavior of generally anthropomorphizing the robot, we observe the
following concrete instances as anthropomorphic effects in this scenario:
talking directly to the robot, or asking it questions (though it doesn't
recognize speech); giving it nick-names or ascribing a gender (\emph{"darling"}
/ \emph{"she"}); praising the robot (\emph{"Good job!"}); using pointing
gestures toward the robot, or try to direct it around with commands; attributing
a personal preference / intention to the robot (\emph{"she loves the hall!"}).

During the following visits at the woman's place, she did not talk to the robot
anymore, and generally behaved in a less anthropomorphizing way. This is why we
suspect the observed effects were due to the novelty of the robot.
